---
title: "HW_7"
author: "Narendra Raj Tiwari"
date: "May 16, 2019"
output: html_document
---


```{r}
library(magrittr)
library(tidyquant)
library(timetk)
library(tibbletime)
library(broom)
library(ggplot2)
library(ggfortify)
library(forecast)
library(Quandl)
library(urca)
library(vars)
library(devtools)
```

a) Getting Data
```{r}
y <- tq_get(("OPHNFB"), get = "economic.data",
           from  = "1947-01-01", to = "2017-12-31")

z <- tq_get(("HOANBS"), get = "economic.data",
           from  = "1947-01-01", to = "2017-12-31")
```


# convert data into tibble
```{r}
y_tbl <-
    y %>%
    as_tibble() %>%
    mutate(yearq = as.yearqtr(date, format = "%Y:%q"))%>%
    dplyr::select(c(yearq,price))


z_tbl<-
  z %>%
    as_tibble() %>%
    mutate(yearq = as.yearqtr(date, format = "%Y:%q"))%>%
    dplyr::select(c(yearq,price))


```

# convert data into ts
```{r}
y_ts <-
    y_tbl %>%
    tk_ts(select = -yearq, start = year(.$yearq[1]), frequency = 4)


z_ts <-
  z_tbl %>%
    tk_ts(select = -yearq, start = year(.$yearq[1]), frequency = 4)
```

b) finding differences#
```{r}
log.y1 <- log(y_ts)
log.y2 <- log(z_ts)

first.diff.y1 <- diff( log(y_ts), differences = 1)
first.diff.y2 <- diff(log(z_ts), differences = 1)
```

checking unit root using ERS test 
```{r}
log.y1 %>% ur.ers(type = "P-test", lag.max = 8, model = "trend") %>% summary()
first.diff.y1 %>% ur.ers(type = "P-test", lag.max = 8, model = "trend") %>% summary()

log.y2 %>% ur.ers(type = "P-test", lag.max = 8, model = "trend") %>% summary()
first.diff.y2  %>% ur.ers(type = "P-test", lag.max = 8, model = "trend") %>% summary()

```

#In first case, we reject the null hypothesis.
#In the second case, we fail to reject the null hypothesis.
#In the third case, we reject the null hypothesis at 1% and 5%.
#In the fourth case, we fail to reject the null hypothesis.

#c
```{r}
x <- cbind(first.diff.y1,first.diff.y2)
var <- VAR(x, ic = "AIC", lag.max = 8, type = "const")
var
summary(var)
```

#d)
```{r}
svar <- BQ(var)
svar
summary(svar)
```

Report and interpretation of the contemporaneous impact and the long run impact matrices#
From the contemporaneous impact matrix, one can easily observe that one standard deviation technologgy shoock 
increases real output per hour 0.0065%.one standard deviation non-technology shock lowers output per hour 
on impact by -0.0032%
Again from the long run impact matrices, a single positive standard deviation technology shock on output per hour
increases by 7.1%, while the non-technology shock is -0.0024.

#e)
```{r}
myIRF.c <- irf(svar, n.head = 12, ci = .9, cumulative = TRUE)
summary(svar)
```

From the contemporaneous impact matrix, one can easily observe that one standard deviation technologgy stock increases real output per hour 0.0065%.While, one standard deviation non-technology shock lowers output per hour on impact by -0.0032%
Also,from the long run impact matrix, a single positive standard deviation technology shock on output per hour  increases it by 0.71% , and non-technology shock decreases output per hour by .24%. 



#f)

```{r, fig.width=10, fig.height=8}
par(mfrow = c(2,2))
plot(myIRF.c, plot.type = "single")
```

#g)


The results in this analysis and Gali(1999) are quite similar though the value and the timing peak are a bit different.

#h)
```{r, fig.width=10, fig.height=8}
plot( fevd(svar, n.ahead=40) ,addbars=10 )
```
Interpretation:
The result shows that larger variation is explained by the internal shocks and 
some part of it is explained by the external shocks.
Moreover, the results holds true for both the long-run and the short-run.
