---
title: "Eco 5316 Time Series Econometrics"
subtitle: Lecture 4 Moving Average (MA) processes
output:
  beamer_presentation:
    includes:
        in_header: lecturesfmt.tex 
    # keep_tex: yes
    highlight: tango
    fonttheme: professionalfonts
fontsize: 9pt
urlcolor: magenta
linkcolor: magenta
editor_options: 
  chunk_output_type: console
---


## MA($q$) model

- let $\{\varepsilon_t\}$ be a white noise time series; process $\{y_t\}$ follows a moving average model of order 1, or MA(1) model, if
$$
    y_t = c_0 + \varepsilon_t + \theta_1 \varepsilon_{t-1}
$$
or, using the lag operator we can write $y_t = c_0 + (1+\theta_1 L) \varepsilon_t$

- the bid-ask bounce in stock trading may introduce an MA(1) structure in a return series

- in general, for $q>0$ a moving average model of order $q$, or MA($q$) model, is given by
$$
    y_t = c_0 + \varepsilon_t + \theta_1 \varepsilon_{t-1} + \ldots + \theta_q \varepsilon_{t-q}
$$
or equivalently using the lag operator $y_t = c_0 + (1+\theta_1 L + \ldots +\theta_q L^q) \varepsilon_t$





## Autocorrelation function for MA($q$) model

- if $q=1$ it can be shown that $\gamma_0=Var(y_t)=(1+\theta_1^2)\sigma_a^2$, and also that $\gamma_1=\theta_1 \sigma_a^2$ and  $\gamma_l=0$ for $l\geq 2$; thus
\begin{equation}
    \rho_1 = \frac{\theta_1}{1+\theta_1^2} \qquad  \rho_l = 0 \text{ for } l > 1
\end{equation}

- theoretical ACF for MA(1) thus cuts off to zero after lag 1

\medskip

- if $q=2$ we have $\gamma_0=Var(y_t)=(1+\theta_1^2+\theta_2^2)\sigma_a^2$, and the theoretical ACF for MA(2) satisfies
\begin{equation}
    \rho_1 = \frac{\theta_1 + \theta_1 \theta_2}{1+\theta_1^2+\theta_2^2} \qquad \rho_2 = \frac{\theta_2}{1+\theta_2^2+\theta_1^2}
    \qquad \rho_l = 0 \text{ for } l > 2
    \label{eq:ACF_MA2}
\end{equation}

- theoretical ACF for MA(1) thus cuts off to zero after lag 2

\medskip

- in general, for MA($q$) model $\gamma_0=Var(y_t)=(1+\theta_1^2+\ldots+\theta_q^2)\sigma_a^2$ 

- theoretical ACF for MA($q$) model satisfies $\rho_l=0$ for $l > q$


## Autocorrelation function for MA($q$) model

- ACF is thus useful for identifying the order of an MA model in the same way the PACF is useful in identifying the order of an AR model

- note that MA($q$) model is always weakly stationary, unlike AR($p$) model which is only weakly stationary if its characteristic roots lie inside the unit circle

- interactive overview of ACF and PACF for simulated MA($q$) models is [here](https://janduras.shinyapps.io/ARMAsim/lec02ARMAsim.Rmd)


## Autocorrelation function for MA($q$) model

```{r, echo=FALSE}
    mymaxlag <- 20
    mynobs <- 500
```

**MA(1) with $\theta_1=0.7$**
```{r, echo=FALSE}
    theta <- c(0.7,0,0)
    y.sim1 <- arima.sim(model=list(order = c(0,0,length(theta)), ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=6}
    par(mfrow=c(2,2), cex=1, mar=c(2,4,0,1))
    plot(0:mymaxlag, ARMAacf(ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(1:mymaxlag, ARMAacf(ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
    acf(y.sim1, type="correlation", lag=mymaxlag, ylab="sample ACF", main="")
    acf(y.sim1, type="partial", lag=mymaxlag, ylab="sample PACF", main="")
```


## Autocorrelation function for MA($q$) model

**MA(1) with $\theta_1=-0.7$**
```{r, echo=FALSE}
    theta <- c(-0.7,0,0)
    y.sim1 <- arima.sim(model=list(order = c(0,0,length(theta)), ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=6}
    par(mfrow=c(2,2), cex=1, mar=c(2,4,1,1))
    plot(0:mymaxlag, ARMAacf(ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(1:mymaxlag, ARMAacf(ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
    acf(y.sim1, type="correlation", lag=mymaxlag, ylab="sample ACF", main="")
    acf(y.sim1, type="partial", lag=mymaxlag, ylab="sample PACF", main="")
```


## Autocorrelation function for MA($q$) model

**MA(2) with $\theta_1=0.2$, $\theta_2=0.7$**
```{r, echo=FALSE}
    theta <- c(0.2,0.7,0)
    y.sim1 <- arima.sim(model=list(order = c(0,0,length(theta)), ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=6}
    par(mfrow=c(2,2), cex=1, mar=c(2,4,1,1))
    plot(0:mymaxlag, ARMAacf(ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(1:mymaxlag, ARMAacf(ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
    acf(y.sim1, type="correlation", lag=mymaxlag, ylab="sample ACF", main="")
    acf(y.sim1, type="partial", lag=mymaxlag, ylab="sample PACF", main="")
```


## Autocorrelation function for MA($q$) model

**MA(2) with $\theta_1=0.2$, $\theta_1=-0.7$**
```{r, echo=FALSE}
    theta <- c(0.2,-0.7,0)
    y.sim1 <- arima.sim(model=list(order = c(0,0,length(theta)), ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=6}
    par(mfrow=c(2,2), cex=1, mar=c(2,4,1,1))
    plot(0:mymaxlag, ARMAacf(ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(1:mymaxlag, ARMAacf(ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
    acf(y.sim1, type="correlation", lag=mymaxlag, ylab="sample ACF", main="")
    acf(y.sim1, type="partial", lag=mymaxlag, ylab="sample PACF", main="")
```


<!--
## Estimation

- MA models are commonly estimated using maximum likelihood

- two approaches: **conditional-likelihood method** and **exact likelihood method**

- conditional maximum likelihood method assumes that initial shocks are zero, $\varepsilon_t=0$ for $t \leq 0$

- exact maximum likelihood method treats initial shocks as additional unknown parameters that are estimated together with $\theta_1, \ldots, \theta_q$

- exact method is more computationally intensive but preferred

- if sample size is large, the two methods yield very similar results

-->
