---
title: "Eco 5316 Time Series Econometrics"
subtitle: Lecture 5 Autoregressive Moving Average (ARMA) processes
output:
  beamer_presentation:
    includes:
        in_header: lecturesfmt.tex 
    # keep_tex: yes
    highlight: tango
    fonttheme: professionalfonts
fontsize: 9pt
urlcolor: magenta
linkcolor: magenta
editor_options: 
  chunk_output_type: console
---


## ARMA($p,q$) model

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 100)
```

```{r echo=FALSE}
knitr::knit_hooks$set(mysize = function(before, options, envir) {
  if (before) 
    return(options$size)
})
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE,  mysize=TRUE, size='\\scriptsize')
```

- AR or MA models may require a high-order model and thus many parameters to adequately describe the dynamic structure of the data

- Autoregressive Moving-Average (ARMA) models allow to overcome this and allow parsimonious model specification with a small number of parameters 



## ARMA($p,q$) model

- suppose that $\{\varepsilon_t\}$ is a white noise, time series process $\{y_t\}$ follows an ARMA(1,1) if
$$
    y_t = \phi_0 + \phi_1 y_{t-1} + \varepsilon_t + \theta_1 \varepsilon_{t-1}
$$
or equivalently, using the lag operator if
$(1-\phi_1 L) y_t = \phi_0 + (1+\theta_1 L) \varepsilon_t$

<!--
If $y_t$ is weakly stationary than  
- unconditional mean: $E(y_t)=\frac{\phi_0}{1-\phi_1}$  
- unconditional variance: $Var(y_t)=\frac{(1+2\phi_1\theta_1+\theta_1^2)\sigma_a^2}{1-\phi_1^2}$  
- ACF: given by $\rho_1 = \phi_1 + \frac{\theta_1 \sigma_a^2}{\gamma_0}$ and $\rho_l=\phi_1 \rho_{l-1}$ for $l>1$, where $\gamma_0$ is the unconditional variance  
- PACF: similar to MA(1), but exponential decay starts with lag 2 not lag 1  
-->

- more generally, time series process $\{y_t\}$ follows an ARMA$(p,q)$ if
$$
    y_t = \phi_0 + \sum_{i=1}^p \phi_i y_{t-i} + \varepsilon_t + \sum_{i=i}^q \theta_q \varepsilon_{t-q}
$$
or, using the lag operator $(1-\phi_1 L-\ldots-\phi_p L^p) y_t = \phi_0 + (1+\theta_1 L+\ldots+\theta_q L^q) \varepsilon_t$



## Autocorrelation function for ARMA($p,q$) model

- recall:  
    - for AR($p$): ACF dies out slowly, PACF drops to zero suddenly after lag $p$  
    - for MA($q$): ACF drops to zero immediately after lag $q$, PACF dies out slowly  

- if neither ACF nor PACF drop to zero abruptly we are dealing with and ARMA model

- in this case both ACF and PACF die out slowly in exponential, oscilating exponential of dampened sine wave pattern

- an overview of ACF and PACF for simulated AR$(p)$, MA$(q)$ and ARMA$(p,q)$ models can be found here:  
[https://janduras.shinyapps.io/ARMAsim/lec02ARMAsim.Rmd](https://janduras.shinyapps.io/ARMAsim/lec02ARMAsim.Rmd)



## Autocorrelation function for ARMA($p,q$) model

\tiny

 process                              ACF                                                   PACF
----------- ------------              ------------------------                              --------------------------
white noise                           $\rho_l=0$ for all $l>0$                              $\phi_{l,l}=0$ for all $l$
   AR(1)    $\phi_1>0$                exponential decay, $\rho_l=\phi_1^l$                  $\phi_{l,l}=\phi_1$, $\phi_{l,l}=0$ for $l>1$ 
            $\phi_1<0$                oscillating decay, $\rho_l=\phi_1^l$                  $\phi_{l,l}=\phi_1$, $\phi_{l,l}=0$ for $l>1$
   AR(2)    $\phi_1^2+4\phi_2>0$      mixture of two exponential decays                     $\phi_{1,1}\neq 0$, $\phi_{2,2}\neq 0$, $\phi_{l,l}=0$ for $l>2$
            $\phi_1^2+4\phi_2<0$      dampened sine wave                                    $\phi_{1,1}\neq 0$, $\phi_{2,2} < 0$, $\phi_{l,l}=0$ for $l>2$
   AR($p$)                            decays toward zero in dampened sine                   $\phi_{l,l} = 0$ for $l>p$
                                      wave pattern or oscillating pattern             
   MA(1)    $\theta_1>0$              $\rho_1>0$, $\rho_l=0$ for all $l>1$                  oscillating decay, $\phi_{1,1}>0$, $\phi_{2,2}<0$, ...
            $\theta_1<0$              $\rho_1<0$, $\rho_l=0$ for all $l>1$                  exponential decay, $\phi_{l,l}<0$ for all $l$
   MA(2)                              $\rho_1 \neq 0, \rho_2 \neq 0$, $\rho_l=0$ for $l>2$  mixture of two direct or oscillatory 
                                                                                            exponential decays, or a dampened wave
   MA($q$)                            $\rho_l = 0$ for $l>q$                                decays toward zero, may oscillate or have
                                                                                            a shape of a dampened sine wave
ARMA(1,1)   $\phi_1>0$, $\theta_1>0$  exponential decay                                     oscilating exponential decay
            $\phi_1>0$, $\theta_1<0$  exponential decay after lag 1                         exponential decay
            $\phi_1<0$, $\theta_1>0$  oscillating exponential decay                         oscillating exponential decay
            $\phi_1<0$, $\theta_1<0$  oscillating exponential decay                         exponential decay
ARMA($p,q$)                           decay (direct or oscillatory) after lag $p$           decay (direct or oscillatory) after lag $q$
                                      or dampened sine wave                                or dampened sine wave



## Autocorrelation function for ARMA($p,q$) model

\scriptsize

```{r, echo=FALSE}
    mymaxlag <- 12
    mynobs <- 500
```

\vspace{+0.15cm}
**ARMA(1,1) with $\phi_1=0.5$, $\theta_1=0.9$**
\vspace{-0.15cm}
```{r, echo=FALSE}
    phi <- c(0.5,0,0)
    theta <- c(0.9,0,0)
    y.sim1 <- arima.sim(model=list(order = c(length(phi),0,length(theta)), ar=phi, ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=1.65}
    par(mfrow=c(1,2), cex=0.5, mar=c(2,4,0,1))
    plot(0:mymaxlag, ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(1:mymaxlag, ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
```

\vspace{-0.15cm}
**ARMA(1,1) with $\phi_1=0.5$, $\theta_1=-0.9$**
\vspace{-0.15cm}
```{r, echo=FALSE}
    phi <- c(0.5,0,0)
    theta <- c(-0.9,0,0)
    y.sim1 <- arima.sim(model=list(order = c(length(phi),0,length(theta)), ar=phi, ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=1.65}
    par(mfrow=c(1,2), cex=0.5, mar=c(2,4,1,1))
    plot(0:mymaxlag, ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(1:mymaxlag, ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
```

\vspace{-0.15cm}
**ARMA(1,1) with $\phi_1=-0.5$, $\theta_1=0.9$**
\vspace{-0.15cm}
```{r, echo=FALSE}
    phi <- c(-0.5,0,0)
    theta <- c(0.9,0,0)
    y.sim1 <- arima.sim(model=list(order = c(length(phi),0,length(theta)), ar=phi, ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=1.65}
    par(mfrow=c(1,2), cex=0.5, mar=c(2,4,1,1))
    plot(0:mymaxlag, ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(1:mymaxlag, ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
```

\vspace{-0.15cm}
**ARMA(1,1) with $\phi_1=-0.5$, $\theta_1=-0.9$**
\vspace{-0.15cm}
```{r, echo=FALSE}
    phi <- c(-0.5,0,0)
    theta <- c(-0.9,0,0)
    y.sim1 <- arima.sim(model=list(order = c(length(phi),0,length(theta)), ar=phi, ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=1.65}
    par(mfrow=c(1,2), cex=0.5, mar=c(2,4,1,1))
    plot(0:mymaxlag, ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(1:mymaxlag, ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
```

$_{}$

<!--

\vspace{-0.15cm}
**ARMA(2,1) with $\phi_1=0.5$, $\phi_1=-0.4$, $\theta_1=0.9$**
\vspace{-0.15cm}

```{r, echo=FALSE}
    phi <- c(0.5,-0.4,0)
    theta <- c(0.9,0,0)
    y.sim1 <- arima.sim(model=list(order = c(length(phi),0,length(theta)), ar=phi, ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=1.5}
    par(mfrow=c(1,2), cex=0.5, mar=c(2,4,0,1))
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
```

\vspace{-0.15cm}
**ARMA(2,1) with $\phi_1=0.5$, $\phi_1=-0.4$, $\theta_1=-0.9$**
\vspace{-0.15cm}
```{r, echo=FALSE}
    phi <- c(0.5,-0.4,0)
    theta <- c(-0.9,0,0)
    y.sim1 <- arima.sim(model=list(order = c(length(phi),0,length(theta)), ar=phi, ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=1.5}
    par(mfrow=c(1,2), cex=0.5, mar=c(2,4,1,1))
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
```

\vspace{-0.15cm}
**ARMA(2,1) with $\phi_1=-0.5$, $\phi_1=0.4$, $\theta_1=0.9$**
\vspace{-0.15cm}
```{r, echo=FALSE}
    phi <- c(-0.5,0.2,0)
    theta <- c(0.5,0,0)
    y.sim1 <- arima.sim(model=list(order = c(length(phi),0,length(theta)), ar=phi, ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=1.5}
    par(mfrow=c(1,2), cex=0.5, mar=c(2,4,1,1))
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
```

\vspace{-0.15cm}
**ARMA(2,1) with $\phi_1=-0.5$, $\phi_1=0.4$ $\theta_1=-0.9$**
\vspace{-0.15cm}
```{r, echo=FALSE}
    phi <- c(-0.5,0.2,0)
    theta <- c(-0.5,0,0)
    y.sim1 <- arima.sim(model=list(order = c(length(phi),0,length(theta)), ar=phi, ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=1.5}
    par(mfrow=c(1,2), cex=0.5, mar=c(2,4,1,1))
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
```



\vspace{-0.15cm}
**ARMA(2,2) with $\phi=(0.9,-0.7)'$, $\theta_1=(-0.5,0.5)'$**
\vspace{-0.15cm}
```{r, echo=FALSE}
    phi <- c(0.9,-0.7,0)
    theta <- c(-0.5,0.5,0)
    y.sim1 <- arima.sim(model=list(order = c(length(phi),0,length(theta)), ar=phi, ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=1.5}
    par(mfrow=c(1,2), cex=0.5, mar=c(2,4,1,1))
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
```

\vspace{-0.15cm}
**ARMA(2,2) with $\phi=(0.2,0.6)'$, $\theta_1=(-0.3,-0.4)'$**
\vspace{-0.15cm}
```{r, echo=FALSE}
    phi <- c(0.2,0.6,0)
    theta <- c(-0.3,0.4,0)
    y.sim1 <- arima.sim(model=list(order = c(length(phi),0,length(theta)), ar=phi, ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=1.5}
    par(mfrow=c(1,2), cex=0.5, mar=c(2,4,0,1))
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
```

**ARMA(2,2) with $\phi=(0.5,0.1)'$, $\theta_1=(0.9,0.9)'$**
```{r, echo=FALSE}
    phi <- c(0.5,0.1,0)
    theta <- c(0.9,0.9,0)
    y.sim1 <- arima.sim(model=list(order = c(length(phi),0,length(theta)), ar=phi, ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=1.5}
    par(mfrow=c(1,2), cex=0.5, mar=c(2,4,1,1))
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
```

\vspace{-0.15cm}
**ARMA(2,2) with $\phi=(0.5,0.1)'$, $\theta_1=(0.1,-0.9)'$**
\vspace{-0.15cm}
```{r, echo=FALSE}
    phi <- c(0.5,0.1,0)
    theta <- c(-0.1,-0.9,0)
    y.sim1 <- arima.sim(model=list(order = c(length(phi),0,length(theta)), ar=phi, ma=theta), n=mynobs)
```

```{r, echo=FALSE, fig.height=1.5}
    par(mfrow=c(1,2), cex=0.5, mar=c(2,4,1,1))
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=FALSE), type='h', xlab="", ylab="theoretical ACF", main="")
    plot(ARMAacf(ar=phi, ma=theta, lag.max=mymaxlag, pacf=TRUE), type='h', xlab="", ylab="theoretical PACF", main="")
```
-->


<!--
## Estimation of AR($p$), MA($q$), and ARMA($p,q$) models

- two likelihood based methods  
    - conditional likelihood method - assumes $\varepsilon_t=0$ for $t\leq 0$  
    - exact likelihood method  - estimates $\varepsilon_t=0$ for $t\leq 0$ together with all other parameters of the model
 
- exact likelihood method is more computationally intensive, but is preferred

- in large samples the difference between the results of the two methods is small
-->


## A Couple of  Notes

- in practice, we rarely find a data series that precisely conforms to a theoretical ACF or PACF

- consequently, there will be some ambiguities when using the Box-Jenkins methodology

- order $(p,q)$ of an ARMA model may depend on the frequency of the series:
    - daily returns of a market index often show some minor serial correlations
    - monthly returns of the index may not contain any significant serial correlation




## Stationarity

- time series $\{y_t\}$ is stationary if it can be represented as a finite order moving average process or a convergent infinite order moving average process

- for an ARMA model to have a convergent MA representation, and thus be stationary, the inverse roots of the polynomial $1-\phi_1 L-\ldots-\phi_p L^p$ must lie inside the unit circle

- for example, for AR(1) the root of $1-\phi_1 x = 0$ is $x = \frac{1}{\phi_1}$ its inverse $\omega = \phi_1$ the condition is thus $|\phi_1|<1$



## Invertibility

- time series $\{y_t\}$ is invertible if it can be represented as a finite order autoregressive process or a convergent infinite order autoregressive process

- for an ARMA model to have a convergent AR representation, and thus be invertible, the inverse roots of the polynomial $1+\theta_1 L+\ldots+\theta_q L^q$ must lie inside the unit circle

- for example, for MA(1) the root of $1+\theta_1 x = 0$ is $x = -\frac{1}{\theta_1}$ its inverse $\omega = -\theta_1$ the condition is thus $|\theta_1|<1$

- to see why this is necessary note that by repeated substitution
\begin{align}
    y_t &= \varepsilon_t + \theta_1 \varepsilon_{t-1} \\
        &= \varepsilon_t + \theta_1 (y_{t-1} - \theta_1 \varepsilon_{t-2}) \\
        &= \varepsilon_t + \theta_1 y_{t-1} - \theta_1^2 (y_{t-2}-\theta_1 \varepsilon_{t-3}) \\
        &= \varepsilon_t + \theta_1 y_{t-1} - \theta_1^2 y_{t-2} + \theta_1^3 (y_{t-3} - \theta_1 \varepsilon_{t-4}) \\
        &= \ldots 
        % &= \varepsilon_t - \sum_{i=1}^\infty (-1)^i \theta_1^i y_{t-i}
\end{align}
we obtain
$$
    \Big(1 + \sum_{i=1}^\infty (-1)^i \theta_1^i L^i \Big) y_t = \varepsilon_t
$$
which requires $|\theta_1|<1$



## Three Representations for an ARMA Model

1. standard representation as ARMA($p,q$)

2. moving average representation of ARMA($p,q$)

3. autoregressive representation of ARMA($p,q$)



## Three Representations for an ARMA Model

1. standard representation as ARMA($p,q$) \medskip

    compact, useful for estimation, and computing forecasts
    $$
     \phi(L) y_t = \phi_0 + \theta(L) \varepsilon_t
    $$
    where $\phi(L) = 1 - \sum_{i=1}^p \phi_i L^i$ and $\theta(L) = 1 + \sum_{i=1}^q \theta_i L^i$ 


## Three Representations for an ARMA Model

2. moving average representation of ARMA($p,q$) \medskip

    if all inverse roots of the equation $\phi(L)=0$ lie inside of the unit circle then $\{y_t\}$ is weakly stationary and can be written as
    $$
        y_t =  \frac{\phi_0 + \theta(L)}{\phi(L)} \varepsilon_t \equiv \frac{\phi_0}{\phi(1)} + \psi(L) \varepsilon_t
    $$
    for AR(1) we for example get 
    $$
        y_t = \frac{1}{1-\phi_1 L} (\phi_0 + \varepsilon_t) = \frac{\phi_0}{1-\phi_1} + \sum_{l=0}^\infty \phi_1^l \varepsilon_{t-l}
    $$
    
    coefficients $\{\psi_i\}$ are referred to as the impulse response function of the ARMA model



## Three Representations for an ARMA Model

3. autoregressive representation of ARMA($p,q$) \medskip

    if all roots of the equation $\theta(L)=0$ lie outside of the unit circle then $\{y_t\}$ is invetible and can be written as
    $$
        \varepsilon_t = \frac{\phi_0 + \phi(L)}{\theta(L)} y_t \equiv \frac{\phi_0}{\theta(1)} + \pi(L) y_t
    $$
    or equivalently
    $$
        y_t = \frac{\phi_0}{1+\theta_1+\ldots+\theta_q} + \sum_{i=l}^\infty \pi_l y_{t-l} + \varepsilon_t
    $$
    
    coefficients $\{\pi_i\}$ are referred to as $\pi$ weights of the ARMA model



## Example: Total Wages and Salaries in Texas

```{r}
library(tidyverse)
library(tidyquant)
library(timetk)
library(ggfortify)
library(forecast)

theme_set(theme_bw())
```

```{r}
# get quarterly Total Wages and Salaries in Texas, Thousands of Dollars, Seasonally Adjusted Annual Rate
wTX_raw <- tq_get("TXWTOT", get = "economic.data", from = "1980-01-01", to = "2018-12-31")

# note that the sample is quite small
wTX_raw

# log change, a stationary transformation
wTX_tbl <- wTX_raw %>%
    rename(wTX = price) %>%
    mutate(dlwTX = log(wTX) - lag(log(wTX))) 
```



## Example: Total Wages and Salaries in Texas

```{r fig.height=5}
wTX_tbl %>%
    tk_xts(date_var = date, select = c("wTX", "dlwTX")) %>%
    autoplot() + 
        labs(x = "", y = "", 
             title = "Wages and Salaries in Texas, Seasonally Adjusted Annual Rate") +
        facet_wrap(~plot_group, ncol = 1, scales = "free", 
                   labeller = labeller(plot_group = c(dlwTX = "Log Change", 
                                                      wTX = "Thousands of Dollars"))) +
        theme(strip.text = element_text(hjust = 0))
```


     
         


## Example: Total Wages and Salaries in Texas

```{r fig.height=2.5}
dlwTX <- wTX_tbl %>%
    filter(!is.na(dlwTX)) %>%
    tk_xts(date_var = date, select = dlwTX) 

nlags <- 24
ggAcf(dlwTX, lag.max = nlags)
ggPacf(dlwTX, lag.max = nlags)
```



## Example: Total Wages and Salaries in Texas

```{r}
m1 <- Arima(dlwTX, order = c(0,0,2))
m1
```



## Example: Total Wages and Salaries in Texas

```{r fig.height=7}
ggtsdiag(m1, gof.lag = nlags)
```



## Example: Total Wages and Salaries in Texas

```{r}
m2 <- Arima(dlwTX, order = c(2,0,0))
m2
```



## Example: Total Wages and Salaries in Texas

```{r fig.height=7}
ggtsdiag(m2, gof.lag = nlags)
```



## Example: Total Wages and Salaries in Texas

```{r fig.height=5}
# z-statistics for coefficients of AR(2) model - phi1 is not signifficant at any level
m2$coef/sqrt(diag(m2$var.coef))
# p values
(1-pnorm(abs(m2$coef)/sqrt(diag(m2$var.coef))))*2

# estimate ARMA model with a restriction on a parameter
m2.rest <- Arima(dlwTX, order = c(2,0,0), fixed = c(0,NA,NA))
m2.rest
```



## Example: Total Wages and Salaries in Texas

```{r fig.height=7}
ggtsdiag(m2.rest, gof.lag = nlags)
```



## Example: Total Wages and Salaries in Texas

```{r fig.height=5}
# find the best ARIMA model based on either AIC, AICc or BIC
m3 <- auto.arima(dlwTX, ic="aicc", seasonal=FALSE, stationary=TRUE)
m3
# by default auto.arima uses stepwise approach and might end up in a "local minimum" like m3 above
m4 <- auto.arima(dlwTX, ic="aicc", seasonal=FALSE, stationary=TRUE, stepwise=FALSE, approximation=FALSE)
m4
```



## Example: Total Wages and Salaries in Texas

```{r fig.height=7}
ggtsdiag(m3, gof.lag = nlags)
```



## Example: Total Wages and Salaries in Texas

```{r fig.height=7}
ggtsdiag(m4, gof.lag = nlags)
```



## Example: Total Wages and Salaries in Texas

```{r fig.height=5}
# check staionarity and invertibility of the estimated model - plot inverse AR and MA roots
autoplot(m4)
```
