---
title: "Eco 5316 Time Series Econometrics"
subtitle: Pipes
output:
  beamer_presentation:
    includes:
        in_header: lecturesfmt.tex 
    fonttheme: professionalfonts
fontsize: 9pt
urlcolor: magenta
linkcolor: magenta
---



```{r echo=FALSE}
knitr::knit_hooks$set(mysize = function(before, options, envir) {
  if (before)
    return(options$size)
})
```

```{r setup, include=FALSE}
options(width=80)
knitr::opts_chunk$set(echo = TRUE,  mysize=TRUE, size='\\small')
```

## Introduction

- pipe operator `%>%` comes from `magrittr` package, see [cran.r-project.org/web/packages/magrittr/vignettes/magrittr.html](https://cran.r-project.org/web/packages/magrittr/vignettes/magrittr.html)

\begin{figure}[h]
    \includegraphics[trim=0cm 0cm 0cm 0cm, clip, height=1in]{figures/magrittr.jpg}
\end{figure}

- [https://en.wikipedia.org/wiki/The_Treachery_of_Images](https://en.wikipedia.org/wiki/The_Treachery_of_Images)

- they help write code that is easier to read and understand

- in particular, pipes are useful for clearly expressing a sequence of operations

- see [http://r4ds.had.co.nz/pipes.html](http://r4ds.had.co.nz/pipes.html) for more details



## Introduction

- a sequence of operations can be carried out in several ways

- you can

    1. save each intermediate step as a new object
    2. overwrite the original object over and over
    3. compose functions
    4. use the pipe


## Example

- how to get a Ph.D. in 5 easy steps

- consider functions

    - `take_courses()`
    - `pass_comps()`
    - `write_papers()`
    - `defend()`
    - `celebrate()`


## Save Intermediate Steps

- how to get a Ph.D.
    ```{r, eval = FALSE}
    you1 <- take_courses(you, n = 20)
    you2 <- pass_comps(you1, max.attempts = 2) 
    you3 <- write_papers(you2, n = 3) 
    you4 <- defend(you3, committee = c("the good", "the bad", "the ugly")) 
    you5 <- celebrate(you4, how.long = "until 3am")
    ```
    
    \normalsize

- downside: forces you to name intermediate elements

- good idea if there are natural names and we want to inspect and further use the intermediate elements

- often there aren't natural names and code becomes cluttered with unintuitive names



## Overwrite the Original

- how to get a Ph.D.
    ```{r, eval = FALSE}
    you <- take_courses(you, n = 20)
    you <- pass_comps(you, max.attempts = 2) 
    you <- write_papers(you, n = 3) 
    you <- defend(you, committee = c("the good", "the bad", "the ugly")) 
    you <- celebrate(you, how.long = "until 3am")
    ```
     
     \normalsize

- downsides: if you make a mistake you'll need to re-run the sequence from the beginning, results for intermediate steps are not available



## Function Composition

- how to get a Ph.D.
    ```{r, eval = FALSE}
    celebrate(
        defend(
            write_papers(
                pass_comps(
                    take_courses(you, n = 20), 
                    max.attempts = 2),
                n = 3), 
            committee = c("the good","the bad","the ugly")), 
        how.long = "until 3am")
    ```
    
    \normalsize

- big disadvantage: hard to see whats going on since
    (1) you have to read from inside-out,  
    (2) function arguments are  spread far apart from the function itself



## Use the Pipe

- how to get a Ph.D.
    ```{r, eval=FALSE}
    you %>%
        take_courses(n = 20) %>%
        pass_comps(max.attempts = 2) %>%
        write_papers(n = 3) %>%
        defend(committee = c("the good","the bad","the ugly")) %>%
        celebrate(how.long = "until 3am")
    ```
    
    \normalsize

- big advantage: easy to follow and understand 



## When Not to Use the Pipe

- pipes are powerful but don't solve every problem

- they are great for rewriting a reasonably short linear sequence of operations

- Hadley Wickham's recommendation: you should reconsider using pipes if

    * your pipes are longer than (say) ten steps - create some intermediate objects with meaningful names (this will make debugging easier and also make it easier to understand code)

    * if there isn't one primary object being transformed and instead you have multiple inputs or outputs

    * if you have a complex dependency structure - pipes are fundamentally linear, expressing complex relationships with them will yield confusing code



## Other Tools From `magrittr`

-  `%$%` is useful for functions that don't take a data frame but rather individual vectors as arguments
    ```{r}
    library(magrittr)
    mtcars %$%
        cor(disp, mpg)
    ```

- `%<>%` operator which allows you to replace code like
    ```{r, eval = FALSE}
    mtcars <- mtcars %>%
        transform(cyl = cyl * 2)
    ```
    with
    ```{r, eval = FALSE}
    mtcars %<>% transform(cyl = cyl * 2)
    ```



## Other Tools From `magrittr`

- `%T>%` works like `%>%` except that it returns the left-hand side instead of the right-hand side
    ```{r fig.height=4}
    rnorm(100) %>%
        matrix(ncol = 2) %T>%
        plot() %>%
        str()
    ```



## Nested Pipes

- note: you can nest pipes

\begin{figure}[h]
    \includegraphics[trim=0cm 0cm 0cm 0cm, clip, height=2in]{figures/yo_dawg_pipes.jpg}
\end{figure}



## Nested Pipes

```{r mysize = TRUE, size = "\\small", eval = FALSE}
# estimate rolling ARIMA model, create 1 period ahead rolling forecasts
results <-
    data.tbl %>%
    as_tbl_time(index = date) %>%
    mutate(arima.model = roll_arima(PAYEMS)) %>%
    filter(!is.na(arima.model)) %>%
    mutate(arima.coefs = map(arima.model, tidy, conf.int = TRUE),
           arima.f = map(arima.model, (. %>% forecast(1) %>% sw_sweep())))
```
