---
title: "Untitled"
author: "Bitaran Jang maden"
date: "May 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
library(magrittr)
library(tidyquant)
library(timetk)
library(tibbletime)
library(broom)
library(ggplot2)
library(ggfortify)
library(forecast)
library(Quandl)
library(urca)
library(vars)
library(devtools)
```
a)

```{r}
MCOILWTICO <- tq_get(("MCOILWTICO"), get = "economic.data",
           from  = "1995-01-01", to = "2017-12-31")

GASREGCOVM<- tq_get(("GASREGCOVM"), get = "economic.data",
           from  = "1995-01-01", to = "2017-12-31")

MCOILWTICO_tbl <-
   MCOILWTICO %>%
    as_tibble() %>%
    mutate(yearq = as.yearqtr(date, format = "%Y:%q"))%>%
    dplyr::select(c(yearq,price))


GASREGCOVM_tbl<-
  GASREGCOVM %>%
    as_tibble() %>%
    mutate(yearq = as.yearqtr(date, format = "%Y:%q"))%>%
    dplyr::select(c(yearq,price))


# convert data into ts
mcoilwtico_ts <-
    MCOILWTICO_tbl %>%
    tk_ts(select = -yearq, start = year(.$yearq[1]), frequency = 12)


gasregcovm_ts <-
    GASREGCOVM_tbl %>%
    tk_ts(select = -yearq, start = year(.$yearq[1]), frequency = 12)

```

## Including Plots

You can also embed plots, for example:

```{r}
loil<-log(mcoilwtico_ts)
lgas<- log(gasregcovm_ts)
ts.plot(loil, lgas, gpars = list(col = c("black", "red")))
```

b) Unit root tests
```{r}
loil.ers <- ur.ers(loil, type="P-test", model="trend")
summary(loil.ers)

lgas.ers <- ur.ers(lgas, type="P-test", model="trend")
summary(lgas.ers)
#we fail to reject null hypothesis for both oil and gas (log transformed). Thus, we conclude that the time series have unit root.

## First Differences of data:
dloil <- diff(loil)
dlgas <- diff(lgas)

dloil.ers <- ur.ers(dloil, type="P-test", model="trend")
summary(dloil.ers)


dlgas.ers <- ur.ers(dlgas, type="P-test", model="trend")
summary(dlgas.ers)

#we  reject the null hypothesis.
## Thus, Crude Oil Prices and Regular Conventional Gas Price are I(1).
```

```{r}
## (c) Determining the Number of Lags:

y <- cbind(loil, lgas)
colnames(y) <- c("log.wti","log.gas")
y <- na.trim(y)
y.VAR.IC <- VARselect(y, type="const")

nlags <- y.VAR.IC$selection["SC(n)"]
nlags

y <- window(y, start=1995+0, end=2010+11/12)
y.CA <- ca.jo(y, ecdet="const", type="trace", K=nlags, spec="transitory")
summary(y.CA)

y.CA <- ca.jo(y, ecdet="const", type="eigen", K=nlags, spec="transitory")
summary(y.CA)

## We reject the null hypothesis. Thus, oil and gas prices are cointegrated.

lttest(y.CA, r=1)

```
The test suggests case 2(restricted constant),
 while the plots from (a) suggest Case 4(restricted trend).



(e) Estimating a Bivariate VEC Model:
```{r}

y.VEC <- cajorls(y.CA, r=1)
y.VEC
```
f) Statistical Significance of alpha1 and alpha2 in the Estimated VEC Model:
```{r}
summary(y.VEC$rlm)
```

We see that alpha1 is not significant, but alpha2 is significant. We also notice that alpha1 >0 and alpha2 >0
Whenever there is a disruption, the signs of the adjustment parameters are consistent with the error correction mechanism.


g)Reestimating the VEC model with a restriction alpha2 = 0:
```{r}

rest.alpha <- matrix(c(1,0), c(2,1))
y.CA.ralpha <- alrtest(y.CA, A=rest.alpha, r=1)
summary(y.CA.ralpha)
```
```
(h) The Intuition for Imposing the Restriction in (f):

 The restriction alpha2=0 means that the adjustment occurs by y1(oil price) only.
 i.e. gas prices cannot affect oil prices.