---
title: "Untitled"
author: "Bikal Koirala"
date: "April 25, 2019"
output: pdf_document
---

```{r}
library(magrittr)
library(tidyquant)
library(timetk)
library(tibbletime)
library(broom)
library(ggplot2)
library(ggfortify)
library(forecast)
library(Quandl)
library(urca)
library(vars)
library(devtools)
```

(a)
```{r}

OPHNFB <- tq_get(("OPHNFB"), get = "economic.data",
           from  = "1947-01-01", to = "2017-12-31")

HOANBS <- tq_get(("HOANBS"), get = "economic.data",
           from  = "1947-01-01", to = "2017-12-31")


# convert data into tibble
OPHNFB_tbl <-
    OPHNFB %>%
    as_tibble() %>%
    mutate(yearq = as.yearqtr(date, format = "%Y:%q")) %>%
  dplyr::select(c(yearq, price))

OPHNFB_ts <-
    OPHNFB_tbl %>%
    tk_ts(select = -yearq, start = year(.$yearq[1]), frequency = 4)

HOANBS_tbl <-
   HOANBS %>%
    as_tibble() %>%
    mutate(yearq = as.yearqtr(date, format = "%Y:%q")) %>%
  dplyr::select(c(yearq, price))


HOANBS_ts <-
    HOANBS_tbl %>%
    tk_ts(select = -yearq, start = year(.$yearq[1]), frequency = 4)



```


(b)
```{r}


    
log.OPHNFB_ts <- log(OPHNFB_ts)
log.HOANBS_ts <- log(HOANBS_ts)

first.diff.y1 <- diff(log(OPHNFB_ts), differences = 1)
first.diff.y2 <- diff(log(HOANBS_ts), differences = 1)


#unit root test
log.OPHNFB_ts %>% ur.ers(type = "P-test", lag.max = 8, model = "trend") %>% summary()

first.diff.y1 %>% ur.ers(type = "P-test", lag.max = 8, model = "trend") %>% summary()

log.HOANBS_ts %>% ur.ers(type = "P-test", lag.max = 8, model = "trend") %>% summary()




first.diff.y2 %>% ur.ers(type = "P-test", lag.max = 8, model = "trend") %>% summary()


```

In the first test (log transformed of Real Output Per Hour of All Persons) we accept the null hypothesis at all level. This means that the log transformed data does  presence unit root.
In the second test regarding first difference of y1 (Real Output Per Hour of All Persons) we reject null hypothesis at 1 percent. Therfore, there is no presence of unit root.
In the third test (log transformed of Hours of All Persons on nonfarm business sector) we accept the null hypothesis at all level. This means that the log transformed data does  presence unit root.
In the fourth test regarding first difference of y2 (Hours of All Persons on nonfarm business sector) we reject null hypothesis at 1 percent. Therfore, there is no presence of unit root.
(c)
```{r}

y <- cbind(first.diff.y1, first.diff.y2)

var.p <- VAR(y, ic = "AIC", lag.max = 8, type = "const")
var.p
summary(var.p)
```
(d)
```{r}
mySVAR <- BQ(var.p)
mySVAR
summary(mySVAR)
```
(e) 

```{r}

myIRF.c <- irf(mySVAR, n.ahead = 12, ci = .9, cumulative = TRUE)

summary(mySVAR)



```

(f)
```{r, fig.width=10, fig.height=8}
par(mfrow=c(2,2))




plot( myIRF.c, plot.type="single" )

```

(g)
```{r}
#As compared to Gai 1999 paper, our result of impulse response function is similar to figure 2 representing to labor and hours. 

```

(h)
```{r, fig.width=10, fig.height=8}
plot( fevd(mySVAR, n.ahead=40) ,addbars=10 )

```