---
title: "HW 7&8"
author: "Talal Alsabhan"
date: "5/16/2019"
output: html_document
---

```{r}
library(timetk)
library(tidyverse)
library(tidyquant)
library(qqplotr)
library(forecast)
library(readr)
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(ggfortify)
library(tsibble)
library(lubridate)
library(zoo)
library(vars)
library(stargazer)
library(listviewer)
library(qqplotr)
library(plotly)
library(wesanderson)
library(listviewer)
```
#HW7
```{r}
#a
OPHNFB_raw <- tq_get("OPHNFB", get = "economic.data", from = "1947-01-01", to = "2017-12-31")
  rY_tbl <-
    OPHNFB_raw %>%
  as_tibble() %>%
  mutate(yearq = as.yearqtr(date, format = "%Y:%q"))%>%
  dplyr::select(c(yearq,price))
  
  rY_ts <-
    rY_tbl %>%
    tk_ts(select = -yearq, start = year(.$yearq[1]), frequency = 4)
  log.rY <- log(rY_ts)
  f.diff.rY <- diff( log(rY_ts), differences = 1)
  
  HOANBS_raw <- tq_get("HOANBS", get = "economic.data", from = "1947-01-01", to = "2017-12-31")  
rH_tbl<-
  HOANBS_raw %>%
    as_tibble() %>%
    mutate(yearq = as.yearqtr(date, format = "%Y:%q"))%>%
    dplyr::select(c(yearq,price))

rH_ts <-
  rH_tbl %>%
  tk_ts(select = -yearq, start = year(.$yearq[1]), frequency = 4)
log.rH <- log(rH_ts)
f.diff.rH <- diff(log(rH_ts), differences = 1)
```
```{r}
#b
log.rY %>% ur.ers( type = c("DF-GLS", "P-test"), model = c( "trend"),
                 lag.max = 8) %>%
  summary()
```
# WE reject the Null
```{r}
#b
f.diff.rY %>% ur.ers( type = c("DF-GLS", "P-test"), model = c( "trend"),
                      lag.max = 4) %>%
  summary()
```
# we fail to reject the Null
```{r}
#b
log.rH %>% ur.ers( type = c("DF-GLS", "P-test"), model = c( "trend"),
                   lag.max = 4) %>%
  summary()
```

# we reject the Null
```{r}
#b
f.diff.rH %>% ur.ers( type = c("DF-GLS", "P-test"), model = c( "trend"),
            lag.max = 4) %>%
  summary()
```
#we fail to reject the Null
```{r}
#c

y <- cbind(f.diff.rY,f.diff.rH)
# estimate VAR(p) using AIC to select p
varp <- VAR(y, ic = "AIC", lag.max = 8, type = "const")
varp
summary(varp)

lmp <- varp$varresult

stargazer(lmp,
          type  ="text", column.labels = rep(colnames(y), 2),
          dep.var.labels.include = FALSE)
# plot residuals and their ACF and PACF
plot(varp)

y<- sweep(y, 2, apply(y,2,mean))
myVar <- VAR(y, ic = "AIC", lag.max = 8)
myVar
summary(myVar)
lmp1 <- myVar$varresult
stargazer(lmp,
          type  ="text", column.labels = rep(colnames(y), 2),
          dep.var.labels.include = FALSE)
```
```{r}
#d
mySVAR <- BQ(myVar)
mySVAR
summary(mySVAR)
```
#e
#The technology shock is increasing by 0.6574%  and decreasing by -0.003188%
#The non-technology shock is increasing by 0.4464% and 0.6029% 
#The long run technology shocks is increasing by 0.7164
```{r}
#f 
irf <- irf(mySVAR, n.ahead=40, ci=.9, cumulative=TRUE)
plot(irf, plot.type="multiple")
```
#e
# I beleive the result here is quiet similar to Gali. However, I got a different peak point for the value and time.
```{r}
#h
varp.fevd <- fevd(mySVAR, n.ahead = 20)
varp.fevd[[1]][c(1,4,8,15),]
varp.fevd[[2]][c(1,4,8,15),]
plot(varp.fevd)
plot(varp.fevd, addbars=8)
```
# in long and short run the fluctuations are explained by internal shochs and a part of it by the external shocks. I #believe the incentive of getting more profit is behind that.





#HW8
```{r}

COP.Data <- tq_get("MCOILWTICO", get = "economic.data", from = "1995-01-01", to = "2017-12-31")
COP_tbl <-
  COP.Data %>%
  as_tibble() %>%
  mutate(yearq = as.yearqtr(date, format = "%Y:%q"))%>%
  dplyr::select(c(yearq,price))

COP_ts <-
  COP_tbl %>%
  tk_ts(select = -yearq, start = year(.$yearq[1]), frequency = 4)
log.COP <- log(COP_ts)
f.diff.COP <- diff( log(COP_ts), differences = 1)


RCGP.Data <- tq_get("GASREGCOVM", get = "economic.data", from = "1995-01-01", to = "2017-12-31")  
RCGP_tbl<-
  RCGP.Data %>%
  as_tibble() %>%
  mutate(yearq = as.yearqtr(date, format = "%Y:%q"))%>%
  dplyr::select(c(yearq,price))

RCGP_ts <-
  RCGP_tbl %>%
  tk_ts(select = -yearq, start = year(.$yearq[1]), frequency = 4)
log.RCGP <- log(RCGP_ts)
f.diff.RCGP <- diff(log(RCGP_ts), differences = 1)

par(mar=c(1,1,1,1))
plot(log.COP, type='l', main="The Log of Oil and Gas Prices from 1995 to 2017"
     ,col="blue", ylim=c(-2, 5),
     xlab= "Year", ylab="log of Price", colour='variable' )
lines(log.RCGP, col="red")

plot(f.diff.COP, type='l', main="Difference of the log of Oil and log of Gas from 1955 to 2017", col="blue",
     ylim=c(-0.4, 0.4))
lines(f.diff.RCGP, col="red")
```
```{r}
#b
log.COP %>% ur.ers( type = c("DF-GLS", "P-test"), model = c( "trend"),
                   lag.max = 8) %>%
  summary()
f.diff.COP %>% ur.ers( type = c("DF-GLS", "P-test"), model = c( "trend"),
                      lag.max = 8) %>%
  summary()

log.COP %>% ur.ers( type = c("DF-GLS", "P-test"), model = c( "trend"),
                   lag.max = 8) %>%
  summary()
f.diff.COP %>% ur.ers( type = c("DF-GLS", "P-test"), model = c( "trend"),
                      lag.max = 8) %>%
  summary()

log.RCGP %>% ur.ers( type = c("DF-GLS", "P-test"), model = c( "trend"),
                   lag.max = 8) %>%
  summary()
f.diff.RCGP %>% ur.ers( type = c("DF-GLS", "P-test"), model = c( "trend"),
                      lag.max = 8) %>%
  summary()
```

```{r}
#c
c <- cbind(f.diff.COP,f.diff.RCGP)
colnames(c) <- c("log.oil","log.gas")
c <- na.trim(c)
VAR.p <- VARselect(c, type="const")
numlags <- VAR.p$selection["SC(n)"]
numlags

c <- window(c, start=1995+0, end=2017+12/12)
c.1 <- ca.jo(c, ecdet="const", type="trace", K=numlags, spec="transitory")
summary(c.1)
c.2 <- ca.jo(c, ecdet="const", type="eigen", K=numlags, spec="transitory")
summary(c.2)

lttest(c.1, r=1)
```

```{r}
#d
restricted <- matrix(c(1,0), c(2,1))
C1.restricted <- alrtest(c.2, A = restricted, r = 1)
VEC.restricted <- cajorls(C1.restricted, r = 1)
summary(VEC.restricted)
```

```{r}
#e
VEC <- cajorls(c.1, r=1)
VEC
```

```{r}
#f
summary(VEC$rlm)
```

# α1 is not statistically significant while α2 is statistically significant. Therefore,there is distrubtion in the long 
#run equilibrium 
```{r}
#g
rest.alpha <- matrix(c(1,0), c(2,1))
c1.rest.alpha <- alrtest(c.1, A=rest.alpha, r=1)
summary(c1.rest.alpha)
```

```{r}
#f
summary(VEC$rlm)
```
# Gas price  has a pure random walk with all the adjustment occuring in oil price when α1 is not zero and α2=0

