---
title: "HW-6"
author: "Narendra"
date: "April 29, 2019"
output: html_document
---

```{r}
library(magrittr)
library(tidyquant)
library(timetk)
library(tibbletime)
library(broom)
library(ggplot2)
library(ggfortify)
library(forecast)
library(vars)
library(stargazer)

```

```{r}
#a
y.data <-
    tq_get(c("GDPC1","GDPDEF"), get = "economic.data",
           from  = "1950-01-01", to = "2017-12-31") %>%
    spread(symbol, price) %>%
    mutate(qtryear = as.yearqtr(date))

s.data <-
    tq_get("^GSPC", get = "stock.prices" ,
           from  = "1950-01-01", to = "2017-12-31") %>%
    mutate(qtryear = as.yearqtr(date)) %>%
    group_by(qtryear) %>%
    summarise(SP500 = mean(adjusted)) %>%
    ungroup()
#b
GDP.data <- full_join(y.data, s.data, by = "qtryear") %>%
  mutate(dlrGDP = 400*(log(GDPC1) - lag(log(GDPC1))))%>%
  mutate (dlrSP500 = 100*((log(SP500) -lag(log(SP500)) )- (log(GDPDEF) - lag(log(GDPDEF)))))%>%
  dplyr::select(qtryear, dlrGDP, dlrSP500) %>%
  na.trim()

GDP.ts <- tk_ts(GDP.data, select = c("dlrGDP","dlrSP500"), start = 1950, frequency = 4)
#c
VARselect(GDP.ts, lag.max = 8, type = "const")

var1 <- VAR(GDP.ts, p = 2, type = "const")
var1
summary(var1)

# estimate VAR(p) using AIC to select p
varp <- VAR(GDP.ts, ic = "AIC", lag.max = 8, type = "const")
varp
summary(varp)


#
mat.r <- matrix(1, nrow = 2, ncol = 5)
mat.r[1, c(1,3)] <- 0
mat.r
varp.r <- restrict(varp, method = "manual", resmat = mat.r)
varp.r
summary(varp.r)
varp.r$restrictions
Acoef(varp.r)

#d

causality(varp, cause = "dlrGDP")
causality(varp, cause = "dlrSP500")
```
# The null hypothes of Graner Casuality is: X does not granger cause Y or the other way around.
#we accept the null hypothesis. That is: dlrSP500  does NOT Granger casue dlrGDP because of higher p-value 
#(0.01>0.05 because we reject the null hypothesis when p<=0.05 from the result table p-value=0.01)
#We conclude that there is no instantaneous causality between growth rate of the U.S real GDP and the inflation
#adjusted annual return of S&P 500.
```{r}
causality(varp, cause = "dlrSP500")
causality(varp.r, cause = "dlrGDP")
```
##We reject the null hypothesis that dlrSP500 does not granger cause dlrGDP,
#thus dlrSP500 granger causes dlrGDP. We conclude that there is instataneous causality between inflation adjusted
#annual returen of S&P 500 and growth rate of the U.S real GDP. Economic intution means that dlrGDP cannot 
#predict dlrSP500 but dlrSP500 can predict dlrGDP1.
```{r}
#e
varp.r.ser <- restrict(varp, method = "ser", thresh = 2.0)
varp.r.ser
summary(varp.r.ser)
varp.r.ser$restrictions
Acoef(varp.r.ser)

# using stargazer package to report results of VAR estimation
lm1 <- var1$varresult
lmp <- varp$varresult

stargazer(lm1,lmp,
          type  ="text", column.labels = rep(colnames(GDP.ts), 2),
          dep.var.labels.include = FALSE)
#f
varp.f <- predict(varp, n.ahead = 4, ci = 0.9 )
plot(varp.f)
fanchart(varp.f)
autoplot(varp.f, is.date = TRUE) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "", y = "", title = "Multistep forecast for House Price Index, quarterly, log change")
#######
var1.irfs <- irf(var1, n.ahead = 10)
varp.irfs <- irf(varp, n.ahead = 10)
# plot IRFs using plot from vars package
par(mfcol=c(2,2), cex = 0.8)
plot(varp.irfs, plot.type = "single", lwd = 2)
# FEVD - based on Choleski decomposition of variance-covariance matrix var(e)
varp.fevd <- fevd(varp, n.ahead = 40)
varp.fevd[[1]][c(1,4,8,15),]
varp.fevd[[2]][c(1,4,8,15),]
plot(varp.fevd)
plot(varp.fevd, addbars=8)

```




