---
title: "HW-6"
author: "Narendra"
date: "April 29, 2019"
output: html_document
---

```{r}
library(magrittr)
library(tidyquant)
library(timetk)
library(tibbletime)
library(broom)
library(ggplot2)
library(ggfortify)
library(forecast)
library(vars)
library(stargazer)

```

```{r}
#a
y.data <-
    tq_get(c("GDPC1","GDPDEF"), get = "economic.data",
           from  = "1950-01-01", to = "2017-12-31") %>%
    spread(symbol, price) %>%
    mutate(qtryear = as.yearqtr(date))

s.data <-
    tq_get("^GSPC", get = "stock.prices" ,
           from  = "1950-01-01", to = "2017-12-31") %>%
    mutate(qtryear = as.yearqtr(date)) %>%
    group_by(qtryear) %>%
    summarise(SP500 = mean(adjusted)) %>%
    ungroup()
#b
GDP.data <- full_join(y.data, s.data, by = "qtryear") %>%
  mutate(dlrGDP = 400*(log(GDPC1) - lag(log(GDPC1))))%>%
  mutate (dlrSP500 = 100*((log(SP500) -lag(log(SP500)) )- (log(GDPDEF) - lag(log(GDPDEF)))))%>%
  dplyr::select(qtryear, dlrGDP, dlrSP500) %>%
  na.trim()

GDP.ts <- tk_ts(GDP.data, select = c("dlrGDP","dlrSP500"), start = 1950, frequency = 4)
#c
VARselect(GDP.ts, lag.max = 8, type = "const")

var1 <- VAR(GDP.ts, p = 2, type = "const")
var1
summary(var1)

# estimate VAR(p) using AIC to select p
varp <- VAR(GDP.ts, ic = "AIC", lag.max = 8, type = "const")
varp
summary(varp)


#
mat.r <- matrix(1, nrow = 2, ncol = 5)
mat.r[1, c(1,3)] <- 0
mat.r
varp.r <- restrict(varp, method = "manual", resmat = mat.r)
varp.r
summary(varp.r)
varp.r$restrictions
Acoef(varp.r)
 For IRFs, from "Orthogonal Impulse Response from dlrGDP" on dlrsp500, it shows the 
shock from the dlrGDP doesn't significantly affect dlrGDP.  On the other hand,  "Orthogonal Impulse Response from dlrsp500" significantly 
affect on dlrGDP as the 95% confidence interval is far away from 0. Hence, there is a positive significant effect of a shock for dlrsp500
 on dlrGDP. Moreover, the positive effect last until somewhere around 4 or 5 quarters that is for about 1 year. Therefore, it doesn't last 
 for quite a long time. For FEVD, the result shows that the shock in dlrGDP does not impact dlrsp500, the shock in dlrsp500 will impact 
 dlrGDP for 40%.
#d

causality(varp, cause = "dlrGDP")
causality(varp, cause = "dlrSP500")
```
# The null hypothes of Graner Casuality is: X does not granger cause Y or the other way around.
#we accept the null hypothesis. That is: dlrSP500  does NOT Granger casue dlrGDP because of higher p-value 
#(0.01>0.05 because we reject the null hypothesis when p<=0.05 from the result table p-value=0.01)
#We conclude that there is no instantaneous causality between growth rate of the U.S real GDP and the inflation
#adjusted annual return of S&P 500.
```{r}
causality(varp, cause = "dlrSP500")
causality(varp.r, cause = "dlrGDP")
```
##We reject the null hypothesis that dlrSP500 does not granger cause dlrGDP,
#thus dlrSP500 granger causes dlrGDP. We conclude that there is instataneous causality between inflation adjusted
#annual returen of S&P 500 and growth rate of the U.S real GDP. Economic intution means that dlrGDP cannot 
#predict dlrSP500 but dlrSP500 can predict dlrGDP1.
```{r}
#e
varp.r.ser <- restrict(varp, method = "ser", thresh = 2.0)
varp.r.ser
summary(varp.r.ser)
varp.r.ser$restrictions
Acoef(varp.r.ser)

# using stargazer package to report results of VAR estimation
lm1 <- var1$varresult
lmp <- varp$varresult

stargazer(lm1,lmp,
          type  ="text", column.labels = rep(colnames(GDP.ts), 2),
          dep.var.labels.include = FALSE)
#f
varp.f <- predict(varp, n.ahead = 4, ci = 0.9 )
plot(varp.f)
fanchart(varp.f)
autoplot(varp.f, is.date = TRUE) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "", y = "", title = "Multistep forecast for House Price Index, quarterly, log change")
#######
var1.irfs <- irf(var1, n.ahead = 10)
varp.irfs <- irf(varp, n.ahead = 10)
# plot IRFs using plot from vars package
par(mfcol=c(2,2), cex = 0.8)
plot(varp.irfs, plot.type = "single", lwd = 2)
# FEVD - based on Choleski decomposition of variance-covariance matrix var(e)
varp.fevd <- fevd(varp, n.ahead = 40)
varp.fevd[[1]][c(1,4,8,15),]
varp.fevd[[2]][c(1,4,8,15),]
plot(varp.fevd)
plot(varp.fevd, addbars=8)

Comments:  From the result, one can easily observe tha tthe forecast for real GDP growth rate in 2019 Q1 is 2.449631, 
the most recent of  GDP estimate of 2019Q1 in "Federal Bank of New York Nowcast" is about 1.43.  
The GDP estimate of 2019Q1 in "GDPNow Federal Bank of Atlanta forecast" is 2.7, and the average figure in
 Wall Street Journal Economic Forecasting Survey is 1.5, the minimum value is 0.5 and the maximum value is 2.94. Hence, 
 the estimated result is in the range of the Wall Street Journal forecasting results, implying the more optimistic result 
 than New York forecast.But it is a bit less optimistic than the Atlanta forecast. 

```




